% include the figures path relative to the master file
\graphicspath{{./content/intro/figures/}}

\section{Attitude estimation}
\label{sec:g-abs-rel}

In this section, we present 2 approaches to estimate the attitude: (i) in the
former method named \emph{absolute rotation}, the absolute rotation and
attitude of the vehicle is estimated under the assumption that the sun position
is known or deduced from time and the GPS location of the vehicle and (ii) in
the latter method named \emph{relative rotation}, the relative rotation of the
vehicle from its initial position is estimated without additional assumption
regarding the sun position. Both approaches required an estimate of the
scattering angle~$\gamma$ which will be presented beforehand.

% In this section, a method is first proposed to estimate the scattering angle~$\gamma$
% without any asumption regarding the constant $\rho_{lmax}$ which can not be estimated in general.
% Then two approaches are considered. The first one, called absolute rotation, makes the asumption
% that the sun position is knwon or can be deduced from time and the GPS location of the vehicle. The second
% called relative rotation enables to estimate the relative rotation of the vehicle from the intial position of the vehicle without any additional assumption regarding the sun position.

\subsection{$\gamma$ estimation}
\label{sec:gamma}

By only measuring the \gls{aop}~$\alpha$ in scattering effects, $\gamma$ needs
to be estimated to get the vector $v$ defined in Eq.\,\eqref{eq:Rwv}. This
equation is valid for all points in sky region. However, only 2 celestial
points are required to estimate $\gamma$ such as:
\begin{equation}
\begin{cases}
R^{t}\cdot s=R_{cp_{1}}\cdot\left[\begin{array}{c}
-\sin\gamma_{1}\sin\alpha_{1}\\
\sin\gamma_{1}\cos\alpha_{1}\\
\cos\gamma_{1}
\end{array}\right]\\
R^{t}\cdot s=R_{cp_{2}}\cdot\left[\begin{array}{c}
-\sin\gamma_{2}\sin\alpha_{2}\\
\sin\gamma_{2}\cos\alpha_{2}\\
\cos\gamma_{2}
\end{array}\right]
\end{cases} \ .
\label{eq:2setseq}
\end{equation}
Using the product of $R_{cp}$ and $R_z(\alpha)$, Eq.\,\eqref{eq:2setseq}
is rewritten as:
% \begin{equation}
% \begin{split}
%   R_{cp_{1}}\cdot
%   \begin{bmatrix}
% \cos\alpha_{1} & -\sin\alpha_{1} & 0\\
% \sin\alpha_{1} & \cos\alpha_{1} & 0\\
% 0 & 0 & 1
% \end{bmatrix}
% \begin{bmatrix}
% 0\\
% \sin\gamma_{1}\\
% \cos\gamma_{1}
% \end{bmatrix}
% \\
% =R_{cp_{2}}\cdot
% \begin{bmatrix}
% \cos\alpha_{2} & -\sin\alpha_{2} & 0\\
% \sin\alpha_{2} & \cos\alpha_{2} & 0\\
% 0 & 0 & 1
% \end{bmatrix}
% \begin{bmatrix}
% 0\\
% \sin\gamma_{2}\\
% \cos\gamma_{2}
% \end{bmatrix}
% \end{split}
% \end{equation}

% leading to:
\begin{equation}
  M_{1}\cdot
  \begin{bmatrix}
0\\
\sin\gamma_{1}\\
\cos\gamma_{1}
\end{bmatrix}
=M_{2}\cdot
\begin{bmatrix}
0\\
\sin\gamma_{2}\\
\cos\gamma_{2}
\end{bmatrix} \ .
\label{eq:2pts}
\end{equation}
By defining the matrix $M$ such that $M=M_{2}^{t}\cdot M_{1}$, $\gamma_1$ and
$\gamma_2$ are found as:
\begin{equation}
\begin{cases}
\gamma_{1}=-\arctan\frac{M_{02}}{M_{01}}\\
\gamma_{2}=-\arctan\frac{M_{20}}{M_{10}}
\end{cases} \ .
\label{eq:gamma-sol}
\end{equation}

The \gls{aop} is $2\pi$ modulus, while the $\gamma$ found in
Eq.\,\eqref{eq:gamma-sol} is $\pi$ modulus leading to two possible solutions
for the vector $v$: 
$\left(\alpha_{1},\gamma_{1}\right)\,\text{and}\,\left(\alpha_{1}+\pi,-\gamma_{1}\right)$.

\subsection{Absolute rotation}
\label{sec:abs-rot}
In order to estimate the absolute rotation and attitude of the \gls{uav}, it is
assumed that: (i) the sun position is known (ii) the vector $v$ is estimated
using the \gls{aop} measures of the sky (2 points) and (iii) the vertical in
the pixel frame is known or a second $w$ is estimated using the \gls{aop} from
horizontal reflected areas. In this study, the vertical in the pixel frame is
assumed to be known.

The aforementioned assumptions lead to the following expression:
\begin{equation}
\begin{cases}
\left[s,z,s\wedge z\right] & =R(t)\cdot\left[v(t),w(t),v(t)\wedge w(t)\right]\\
 & =R_{wv}(t)\cdot R_{vc}\cdot\left[v(t),w(t),v(t)\wedge w(t)\right]
\end{cases} \ .
\label{eq:linear_equation}
\end{equation}
\noindent where $z$ is the vertical in world frame ($[0, 0, 1]$) and $t$ is the
time instance.

Solving Eq.\,\eqref{eq:linear_equation} enables to get $R_{wv}(t)$. However, due to
$\gamma$ ambiguities, $v$ and therefore $R_{wv}$ have 2 solutions. At each
iteration, the rotation $R_{wv}$ selected is the one the closest from previous
rotation, assuming that the motion between two frames is smoothed.

\subsection{Relative rotation}
\label{sec:rel-rot}

The relative rotation is estimated between two time stamps ($t_1$,
$t_2$). Let $v(t_1)$ and $v(t_2)$ referring to $v_1$ and $v_2$ to simplify the
expression. Therefore, Eq.\,\eqref{eq:linear_equation} becomes:

\begin{equation}
\begin{cases}
\left[s,z,s\wedge z\right] & =R_{wv1}\cdot R_{vc}\cdot\left[v_{1},w_{1},v_{1}\wedge w_{1}\right]\\
\left[s,z,s\wedge z\right] & =R_{wv2}\cdot R_{vc}\cdot\left[v_{2},w_{2},v_{2}\wedge w_{2}\right]
\end{cases}.
\label{eq:rel-linear_equation}
\end{equation}
Leading to:
\begin{equation}
  \begin{split}
R_{wv2}=R_{wv1}\cdot R_{vc}\cdot\left[v_{1},w_{1},v_{1}\wedge
  w_{1}\right]\cdot \\
\left[v_{2},w_{2},v_{2}\wedge w_{2}\right]^{-1} \cdot R_{vc}^{T} \ ,
\label{eq:relative_equation}\end{split}
\end{equation}

Using the above equation, the relative rotation $R_{v_{1}v_{2}}$ is equal to:
%$=R_{wv1}^{T}\cdot R_{wv2}$ is therefore equal to:
\begin{equation}
\begin{split}
  R_{v_{1}v_{2}} &= \\
   & R_{vc}\cdot\left[v_{1},w_{1},v_{1}\wedge
  w_{1}\right]\cdot\left[v_{2},w_{2},v_{2}\wedge w_{2}\right]^{-1}\cdot
R_{vc}^{t}
\end{split} \ .
\label{eq:final-relative}
\end{equation}

As previously explained, only 2 points are required to compute the scattering
angle $\gamma$. In practice (see Sect.\,\ref{sec:exp-res}), more points can be
used which leading to a more robust estimate of the vehicle rotation.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
